---
title: "Burns Sepsis"
author: "Sebastian"
date: "2025-07-15"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r README}

# KCH UKL Burns Sepsis study
```

## KCH UKL Burns Sepsis 

```{r, echo=FALSE}
# Loading packages
library(readxl)
library(writexl)
library(gtsummary)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
library(broom)
library(forestplot)
library(purrr)
```

```{r data}
# loading data frames

#Clinical data
Burns_Sepsis <- read_excel("~/UKL/Malawi ABS Projekt/R Malawi AMS/Burns_Sepsis_09.26.xlsx")

#clinical data raw
CRF1_5_9.26 <- read_excel("~/UKL/Malawi ABS Projekt/R Malawi AMS/CRF1_5_09.26_deidentified.xlsx")

#Microbiology data
Microbiology <- read_excel("~/UKL/Malawi ABS Projekt/R Malawi AMS/Microbiology_09.26.xlsx")
```

### Variable Exploration

```{r}
hist(Burns_Sepsis$age_years)
```

```{r}
ggplot(data = Burns_Sepsis, aes(x = time_to_admission)) +
  geom_histogram(bins = 30)
```

```{r}
Burns_Sepsis %>%
  ggplot()+
  geom_histogram(mapping = aes(x = duration_of_stay), bins = 20)
```

```{r}
hist(Burns_Sepsis$TBSA_mean)
```

Numerical variables are skewed to the right and not normally distributed. Assumption for e.g. t-test, linear regression not met. Mean not a good summary.

### Variable transformation

```{r, echo=FALSE}

# age2 variable with 2 categories 
# [3y because it creates the most evenly sized groups]
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(age2 = as.factor(case_when(
    age_years <= 3 ~ "<=3",
    age_years >3 ~ ">3",
    TRUE ~ NA_character_
  )))

# reference: age >3
Burns_Sepsis$age2 <- relevel(Burns_Sepsis$age2, ref = ">3")

# residence 
# setting reference level
Burns_Sepsis$residence <- relevel(factor(Burns_Sepsis$residence), ref =
                                    "outside_Lilongwe")

# Underweight status 
# [no sig. difference between underweight and severe underweight --> combined]
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(underweight_combined = as.factor(case_when(
    underweight_status %in% c("Severely_underweight", "Underweight") ~ "Underweight",
    underweight_status == "Not_underweight" ~ "Not_underweight",
    TRUE ~ NA_character_
  )))

# setting reference level
Burns_Sepsis$underweight_combined <- relevel(Burns_Sepsis$underweight_combined, ref = "Not_underweight")

# TBSA2 categories
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(TBSA2 = as.factor(case_when(
    TBSA_mean <= 15 ~ "<=15",
    TBSA_mean >15 ~ ">15",
    TRUE ~ NA_character_
  )))

# setting reference level
Burns_Sepsis$TBSA2 <- relevel(Burns_Sepsis$TBSA2, ref = "<=15")

# Type 3
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(type3 = as.factor(case_when(
    type %in% c("Electric", "Other") ~ "Other",
    type %in% c("Scald_water", "Scald_other") ~ "Scald",
    type == "Flame" ~ "Flame",
    TRUE ~ NA_character_
  )))

# setting reference level
Burns_Sepsis$type3 <- relevel(Burns_Sepsis$type3, ref = "Scald")

# Degree
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(degree = as.factor(case_when(
    degree_highest == 1 ~ "1",
    degree_highest %in% c(2,3) ~ "2a-b",
    degree_highest %in% c(4,5) ~ "3-4",
    TRUE ~ NA_character_
  )))
# reference level
Burns_Sepsis$degree <- relevel(Burns_Sepsis$degree, ref = "1")

# time to admission
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(Time_to_admission = as.factor(case_when(
    time_to_admission <= 1 ~ "1d",
    time_to_admission >1 ~ ">1d",
    TRUE ~ NA_character_
  )))
# reference level
Burns_Sepsis$Time_to_admission <- relevel(Burns_Sepsis$Time_to_admission, 
                                            ref = "1d")

# Topical variable
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(Topical_combined = as.factor(case_when(
    Topical_Any == 1 & Topical_Eggs == 0 & Topical_Herbs == 0 & Topical_Flour == 0 ~ "Other",
    Topical_Herbs == 1 ~ "Herbs",
    Topical_Eggs == 1 ~ "Eggs",
    Topical_Flour == 1 ~ "Flour",
    Topical_Any == 0 ~ "None",
    TRUE ~ NA_character_
    )))
#reference level
Burns_Sepsis$Topical_combined <- relevel(Burns_Sepsis$Topical_combined, ref = "None")

# Burns_Sepsis %>%
#   select(Topical_combined) %>%
#   tbl_summary()

#Topical2 (any vs. none)
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(Topical2 = as.factor(case_when(
    Topical_Any == 1 ~ "Any",
    Topical_Any == 0 ~ "None",
    TRUE ~ NA_character_
    )))

#reference level
Burns_Sepsis$Topical2 <- relevel(Burns_Sepsis$Topical2, ref = "None")

# duration3
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(duration3 = as.factor(case_when(
    duration_of_stay <= 7 ~ "<=7",
    duration_of_stay > 7 & duration_of_stay <= 14 ~ ">7-14",
    duration_of_stay > 14 ~ ">14",
    TRUE ~ NA_character_
  )))
# reference duration
Burns_Sepsis$duration3 <- relevel(Burns_Sepsis$duration3, ref = "<=7")


#Factorizing & labelling other variables
Burns_Sepsis <- Burns_Sepsis %>% 
  mutate(
    age = age_years,
    sex = factor(crf1_sex, labels = c("female", "male")),
    underweight_status = factor(underweight_status),
    bmi_age = factor(bmi_age),
    residence = factor(crf1_residence.x, labels = c("outside_Lilongwe",
                                                       "Lilongwe")),
    inhalation_injury = factor(crf1_inhalation_injury, labels = c("no", "yes")),
    Topical_Eggs = factor(Topical_Eggs),
    Topical_Flour = factor(Topical_Flour),
    Topical_Herbs = factor(Topical_Herbs),
    Cleaning_Water = factor(Cleaning_Water),
    Pseudomonas = factor(Pseudomonas),
    MDR = factor(MDR),
    bedsheets = factor(crf1_bedsheets, labels = c("not provided", "provided")),
    Escharotomy = factor(Surg_Escharotomy),
    Debridement = factor(Surg_Debridement),
    Split_skin_graft = factor(Surg_Split_skin_graft),
    urban_vs_rural = factor(urban_vs_rural)
  )

#reference level - Bedsheets 
Burns_Sepsis$bedsheets <- relevel(factor(Burns_Sepsis$bedsheets), ref = "provided")
```

### Descriptive findings

#### Participants

```{r}
# Potentially eligible:
  # 210 paediatric patients admitted to the unit during the observation period

# Screened: 105

# Included
Burns_Sepsis %>% 
  filter(!is.na(study_id)) %>% 
  n_distinct()

# Completed follow up
Burns_Sepsis %>% 
  filter(!is.na(crf5_assessment_date)) %>% 
  count()

# Sufficient data for sepsis endpoint 
Burns_Sepsis %>% 
  filter(!is.na(crf5_sepsis_corr)) %>% 
  count()
```

-   210 eligible patients were admitted during study the period
-   105 patients were screened
-   104 patients were included in the study
-   103 patients completed follow up
-   93 patients with sufficient vital sign data to assess sepsis endpoint

#### Baseline characteristics

```{r baseline_Table_1}
# Patient characterstics (Table 1)

Burns_Sepsis %>%
  select(age, sex, underweight_status, bmi_age, TBSA_mean, degree_highest, type,
         inhalation_injury, crf1_surgical) %>% 
  tbl_summary()
```

#### Previous facility

```{r referral_tableS1}

Burns_Sepsis %>%   
  select(starts_with("Facility_")) %>%    
  tbl_summary()
```

```{r Comorbidities}

sum(CRF1_5_9.26$crf1_comorb, na.rm = TRUE) #9
sum(CRF1_5_9.26$crf1_comorb_list___1, na.rm = TRUE) #Epilepy 5
sum(CRF1_5_9.26$crf1_comorb_list___2, na.rm = TRUE) #Psych.disorder 0
sum(CRF1_5_9.26$crf1_comorb_list___3, na.rm = TRUE) #HIV 0
sum(CRF1_5_9.26$crf1_comorb_list___4, na.rm = TRUE) #DM T1 0
sum(CRF1_5_9.26$crf1_comorb_list___5, na.rm = TRUE) #Malnutrition 2
sum(CRF1_5_9.26$crf1_comorb_list___6, na.rm = TRUE) #Other 3 --> Malaria, Cerebral Palsy, Asthma
```

#### Topical applications & wound cleaning

```{r topical_cleaning}
Burns_Sepsis %>%
  select(starts_with("Topical_"), Cleaning_Water) %>% 
  tbl_summary()
```

Other = “local medicine”, “traditional meds”, “brick powder”, “soda”, “lotion”, “black stuff”; and "unknown substances".

### **AMS rounds**

```{r AMS_round_desriptive}
#Nr. of AMS rounds: 35
CRF1_5_9.26 %>%
  filter(!is.na(crf4_assessment_date)) %>% 
  select(crf4_assessment_date) %>% 
  n_distinct()

#Nr. of cases discussions: 109
CRF1_5_9.26 %>%
  filter(!is.na(crf4_assessment_date)) %>%
  select(crf4_assessment_date) %>% 
  nrow()

#Nr. of individual patients discussed: 57
CRF1_5_9.26 %>%
  filter(!is.na(crf4_entry_date)) %>%
  distinct(record_id) %>%
  n_distinct()
```

35 AMS ward rounds were conducted in which 109 case discussions were held including 57 patients.

### Antibiotic therapies

```{r Abx_therapy}
# patients with antimicrobial therapy at time of AMS round: 59
CRF1_5_9.26 %>%
  filter(crf4_antimicrobial == 1) %>%
  nrow()

# In percent of cases discussed
59/109  #54%

# Patients without antimicrobial therapy at time of AMS round: 50
CRF1_5_9.26 %>%
  filter(crf4_antimicrobial == 0) %>%
  nrow()

# Antimicrobial therapies observed during AMS rounds
CRF1_5_9.26 %>%
  select(record_id, crf4_entry_date, crf4_antimicrobial_substance1, 
         crf4_antimicrobial_substance2, crf4_antimicrobial_substance3) %>%
  filter(!is.na(crf4_antimicrobial_substance1)) %>%
  rowwise() %>%
  mutate(substances = list(sort(na.omit(c(crf4_antimicrobial_substance1, 
                                          crf4_antimicrobial_substance2, 
                                          crf4_antimicrobial_substance3))))) %>%
  ungroup() %>%
  group_by(substances) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count)) %>%
  view()
```

At the time of review, 54% (59/109) of cases were on antimicrobials, most frequently ceftriaxone (65%; 31/48).

**AMS recommendations**

```{r AMS_recommendations}

#Descriptive Table
CRF1_5_9.26 %>%
  select(starts_with("AMS_", ignore.case = TRUE)) %>% 
  tbl_summary()

#Any? corrected  n = 109
CRF1_5_9.26 %>% 
  count(AMS_Any_new)
57/109     #52%

#Other recommendation
CRF1_5_9.26 %>% 
  select(crf4_ams_intervention_list_oth) %>% 
  tbl_summary()

#Nr. of follow up discussions: 53
CRF1_5_9.26 %>% 
  filter(!is.na(crf4_ams)) %>% 
  count(crf4_ams)  #included in previous AMS round

#Implementation of AMS recommendations?
CRF1_5_9.26 %>% count(crf4_ams_impl)
#yes: 37; partial: 3

37/53 #70%
```

The AMS team issued recommendations in 52% (57/109) of cases, with microbiological testing being the most common recommendation (40%; 23/57). In 70% (37/53) of follow-up discussions, the previous AMS recommendations had been implemented.

#### Clinical outcomes

```{r duration}
# Duration of stay 
Burns_Sepsis %>%
  select(duration_of_stay) %>%
  tbl_summary()
```

median length of stay: 9 days (IQR 5-17)

```{r clinical_outcomes}
# Clinical outcomes
Burns_Sepsis %>%
  select(wound_infection_any, crf5_sepsis_corr, Death_corr) %>% 
  tbl_summary() %>% 
  add_ci()
```

Wound infection: 58 (56%, 95% CI: 46% to 65%) \
Sepsis: 35 (38%, 95% CI: 28% to 48%)\
Death: 22 (21%, 95% CI: 14% to 31%)

#### Microbiological results

```{r Nr_swabs}
#Nr. of Blood cultures and wound swabs: 44
Microbiology %>% 
  count(`Specimen Type`)

#Nr. of individual patients with wound swabs: 30
Microbiology %>% 
  filter(`Specimen Type` == "Wound Swab") %>%
  distinct(study_id) %>% 
  n_distinct()

#Nr. of wound swabs with growth: 39
Microbiology %>% 
  count(Woundswab_growth)

#Nr. of patients with wound swabs growth: 27
Microbiology %>% 
  filter(Woundswab_growth == 1) %>% 
  distinct(study_id) %>% 
  n_distinct()
```

44 deep wound swabs were taken from 30 study participants of which 39 (89%) revealed bacterial growth.

```{r Nr_blood_cultures}

#Nr. of Blood cultures and wound swabs: 13
Microbiology %>% 
  count(`Specimen Type`)

#Nr. of individual patients with blood cultures: 10
Microbiology %>% 
  filter(`Specimen Type` == "Blood") %>% 
  distinct(study_id) %>% 
  n_distinct()

#Nr. with blood culture growth: 4
Microbiology %>% 
  count(Bloodculture_growth)

#Nr. of patients with blood culture growth: 4
Microbiology %>% 
  filter(Bloodculture_growth == 1) %>% 
  distinct(study_id) %>% 
  n_distinct()

#Nr. organisms identified in blood cultures
Microbiology %>% 
  filter(Bloodculture_growth == 1) %>%
  count(Organism)
```

A total of 13 blood culture pairs were taken from 10 study participants, of which 4 showed bacterial growth: two had a bacteraemia with *S. aureus*, one with *A. baumannii* and one with *Klebsiella pneumoniae*.

```{r Organisms_total}

Microbiology %>% 
  select(Organism) %>% 
  filter(Organism != "No growth") %>% 
  tbl_summary()

#gram negative: 32
4+4+1+4+1+18

#gram positive: 11
10+1

#gram negative percent: 74%
32/43
```

```{r Organisms_wound_swabs}

Microbiology %>% 
  filter(Woundswab_growth == 1) %>% 
  select(Organism) %>% 
  tbl_summary()

#gram negative: 30
3+4+1+3+1+18

#gram positive: 9
8+1

#gram negative percent: 77%
30/39
```

The majority of identified pathogens were Gram-negative bacteria (30, 77%). The most frequently isolated Gram-negative pathogen was *P. aeruginosa* (18, 46%), followed by *Escherichia* (*E*.) *coli* and *Klebsiella* spp. each with 4 isolates and *A.* *baumannii* with three isolates.

Among Gram-positive pathogens, *S. aureus* was identified most frequently (8, 20%), while streptococci were detected only once. Coagulase-negative staphylococci and enterococci were not identified in any wound swabs.

```{r AST_carbapenem_resistance}

# Pseudomonas Carbapenem resistance
Microbiology %>% 
  filter(Organism == "Pseudomonas aeruginosa") %>% 
  select(Meropenam) %>% 
  tbl_summary()
  
# Acinetobacter Carbapenem resistance
Microbiology %>% 
  filter(Organism == "Acinetobacter baumanni") %>% 
  select(Meropenam) %>% 
  tbl_summary()
```

*P. aeruginosa* and *A. baumannii* were resistant to carbapenems in 22% (4/18) and 50% (2/4) of cases, respectively.

```{r MDR}

#MDR Acinetobacter (4/4)
Microbiology %>% 
  filter(Organism == "Acinetobacter baumanni") %>% 
  select(MDR) %>% 
  tbl_summary()

#MDR Klebsiella spp. (3/5)
Microbiology %>% 
  filter(grepl("Klebsiella", Organism, ignore.case = TRUE)) %>% 
  select(MDR) %>% 
  tbl_summary()

#MDR Klebsiella pneumoniae (3/4)
Microbiology %>% 
  filter(Organism == "Klebsiella pneumoniae") %>% 
  select(MDR) %>% 
  tbl_summary()

#MDR Pseudomonas (3/13)
Microbiology %>% 
  filter(Organism == "Pseudomonas aeruginosa") %>% 
  select(MDR) %>% 
  tbl_summary()

#MDR all (16/38)
Microbiology %>% 
  select(MDR) %>% 
  tbl_summary()
```

MDR phenotypes were found in all (4/4) *A. baumannii* isolates, 75% (of *Klebsiella* spp. and 23% (3/13) of *Pseudomonas* isolates.

Among all isolates with sufficient AST, 42% (16/38) were MDR.

```{r save_Microbiology}

write_xlsx(Microbiology, "~/UKL/Malawi ABS Projekt/R Malawi AMS/Microbiology_2026_09_22.xlsx")
```

#### Treatment

```{r surgeries_bedsheets}

Burns_Sepsis %>% 
  select(bedsheets, Debridement, Split_skin_graft) %>% 
  tbl_summary()
```

```{r surgeries}
Burns_Sepsis %>%
  select(Surg_Escharotomy, Surg_Debridement, Surg_Split_skin_graft,
         Surg_Fascitotomy, Surg_Amputation, Surg_Other_surgery) %>% 
  tbl_summary()
```

Upon admission to the unit, 53 (51%) study participants received fresh bedsheets. Surgical debridement was done in 18 (17%) cases, while 16 (15%) study participants received skin grafting.

[All antibiotic treatment]{.underline}

```{r antibiotics_all}
#Antiinfecte treatments & combinations (total)
CRF1_5_9.26 %>%
  select(record_id, crf4_entry_date, crf4_antimicrobial_substance1, 
         crf4_antimicrobial_substance2, crf4_antimicrobial_substance3) %>%
  filter(!is.na(crf4_antimicrobial_substance1)) %>%
  rowwise() %>%
  mutate(substances = list(sort(na.omit(c(crf4_antimicrobial_substance1, 
                                          crf4_antimicrobial_substance2, 
                                          crf4_antimicrobial_substance3))))) %>%
  ungroup() %>%
  group_by(substances) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count)) %>% 
  view()
```

[Empirical treatment]{.underline}

The most commonly observed empirical antibiotic therapies were ceftriaxone (12; 43%) and ceftriaxone in combination with metronidazole (10; 36%).

```{r antibiotics_empirical}
#Empirical therapies
# (CRF4: ABx; CRF4: Microbiological testing: "no")
CRF1_5_9.26 %>%
  filter(!is.na(crf4_antimicrobial_substance1) & crf4_microbial_test ==0) %>%
  rowwise() %>%
  mutate(substances = list(sort(na.omit(c(crf4_antimicrobial_substance1, 
                                          crf4_antimicrobial_substance2, 
                                          crf4_antimicrobial_substance3))))) %>%
  ungroup() %>%
  group_by(substances) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count)) %>%
  view()
```

[Treatment after microbial testing]{.underline}

```{r antibiotics_targeted}
#(CRF4: ABx; CRF4: Mibi: "yes")

CRF1_5_9.26 %>%
  filter(!is.na(crf4_antimicrobial_substance1) & crf4_microbial_test ==1) %>%
  rowwise() %>%
  mutate(substances = list(sort(na.omit(c(crf4_antimicrobial_substance1, 
                                          crf4_antimicrobial_substance2, 
                                          crf4_antimicrobial_substance3))))) %>%
  ungroup() %>%
  group_by(substances) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count)) %>%
  view()
```

## Determinants of wound infection, sepsis & death

```{r age}

t.test(age_years ~ Death_corr, data = CRF_stat, na.action = na.omit) 
```

There was no significant difference in outcomes between male and female participants. However, age showed an inverse relationship with mortality: children who died were younger on average than those who survived (2·8 vs. 4·0 years).

```{r admission_time}

t.test(Burns_Sepsis$Death_corr, Burns_Sepsis$time_to_admission)
```

In fact, deceased participants were admitted significantly earlier on average than survivors (mean 0·2 vs. 9 days; t-test: p = 0.006).

### **Univariate log regression  & forrest plots**

```{r exposures}
# exposures included in forrest plots


exposures <- c("sex", "age2", "residence", "underweight_combined", 
               "TBSA2", "type3", "degree", "Time_to_admission",
               "Topical2", "Cleaning_Water", 
               "duration3", "MDR", "bedsheets")

```

**Univariate log regression wound infection**

```{r FP_WI_table}

# Creating DF for forrest plot Wound infection

# descriptive table
FP_table_WI <- Burns_Sepsis %>%
  select(all_of(exposures), wound_infection_any) %>%
  pivot_longer(cols = all_of(exposures), names_to = "variable", values_to = "category") %>%
  mutate(variable = factor(variable, levels = exposures)) %>%   #keeping order defined in "exposures"
  group_by(variable, category) %>%
  summarise(
    n = n(),
    wound_infection = sum(wound_infection_any == 1, na.rm = TRUE),
    wound_infection_pct = round(100 * wound_infection / n, 1),
    .groups = "drop"
  ) %>% 
  arrange(variable) #keeping order of variables as given in exposures

print(FP_table_WI, n = 100)

write_xlsx(FP_table_WI, "~/UKL/Malawi ABS Projekt/R Malawi AMS/discriptive_WI.xlsx")

# OR results table
results_WI <- exposures %>%
  set_names() %>%
  map_df(~ glm(as.formula(paste("wound_infection_any ~", .x)), data = Burns_Sepsis,
               family = binomial("logit")) %>% 
           tidy(exponentiate = TRUE, conf.int = TRUE), .id = "exposure") %>%
  filter(term != "(Intercept)")

# Keeping order ... 
results_WI <- results_WI %>%
  mutate(
    term = factor(term, levels = unique(results_sepsis$term)),   
    exposures = factor(exposure, levels = exposures))

print(results_WI, n = 100)

results_WI <- results_WI %>% 
  select(term, estimate, conf.low, conf.high)

# Saving data
write_xlsx(results_WI, "~/UKL/Malawi ABS Projekt/R Malawi AMS/results_WI.xlsx")

# ordering manually in excel

# reading df
fp_WI_data <- read_excel("~/UKL/Malawi ABS Projekt/R Malawi AMS/Forrest_plot_data_wi.xlsx")

OR_CI_wi <- ifelse(
  fp_WI_data$estimate == 1.0,
  "1.0 (Ref)",
  sprintf("%.1f (%.1f–%.1f)", fp_WI_data$estimate, fp_WI_data$conf.low, fp_WI_data$conf.high)
)

FP_WI_df <- data.frame(
  Variable = c("Characteristic", fp_WI_data$category),
  N = c("N", fp_WI_data$n),
  Wound_infection = c("Wound Infection", fp_WI_data$wound_infection),
  Wound_infection_percent = c("%", fp_WI_data$wound_infection_pct),
  OR_CI = c("OR (95% CI)", OR_CI_wi),
  stringsAsFactors = FALSE
)

mean_wi <- c(NA, fp_WI_data$estimate)
lower_wi <- c(NA, fp_WI_data$conf.low)
upper_wi <- c(NA, fp_WI_data$conf.high)

table_text_wi <- as.matrix(FP_WI_df)
```

**Wound infection - Forrest plot**

```{r Figure2_FP_WI_plot}

# Plotting
forestplot(
  labeltext = table_text_wi,
  mean = mean_wi,
  lower = lower_wi,
  upper = upper_wi,  
  is.summary = c(TRUE, rep(FALSE, nrow(table_text_wi) -1)),
  zero = 1,
  xlog = TRUE,
  xlim = c(0.01, 50),
  xticks = c(0.01, 0.1, 0.5, 1, 2, 5, 10, 20, 50),  
  colgap = unit(3, "mm"),
  col = fpColors(box = "black", lines = "gray", zero = "black"),
  boxsize = 0.2,
  txt_gp = fpTxtGp(
    summary = gpar(cex = 1),
    label = gpar(cex = 0.8),
    ticks = gpar(cex = 0.8)
  ),
  lineheight = unit(0.5, "cm"),
  line.margin = 0.1,
  graph.pos = 5
)
```

**Univariate log regression sepsis - Table**

```{r FP_sepsis_table}

# Creating DF for forrest plot Sepsis

Burns_Sepsis_corr <- Burns_Sepsis %>% 
  filter(!is.na(crf5_sepsis_corr)) #filtering sepsis status != NA
  
# descriptive table
FP_table_sepsis <- Burns_Sepsis_corr %>%
  select(all_of(exposures), crf5_sepsis_corr) %>%
  pivot_longer(cols = all_of(exposures), names_to = "variable", values_to = "category") %>%
  mutate(variable = factor(variable, levels = exposures)) %>%   #keeping order defined in "exposures"
  group_by(variable, category) %>%
  summarise(
    n = n(),
    sepsis = sum(crf5_sepsis_corr == 1, na.rm = TRUE),
    sepsis_pct = round(100 * sepsis / n, 1),
    .groups = "drop"
  ) %>% 
  arrange(variable) #keeping order of variables as given in exposures

print(FP_table_sepsis, n = 100)

write_xlsx(FP_table_sepsis, "~/UKL/Malawi ABS Projekt/R Malawi AMS/discriptive_sepsis_corr.xlsx")

# OR results table
results_sepsis <- exposures %>%
  set_names() %>%
  map_df(~ glm(as.formula(paste("crf5_sepsis_corr ~", .x)), data = Burns_Sepsis_corr, 
               family = binomial("logit")) %>% 
           tidy(exponentiate = TRUE, conf.int = TRUE), .id = "exposure") %>%
  filter(term != "(Intercept)")

# Keeping order ... 
results_sepsis <- results_sepsis %>%
  mutate(
    term = factor(term, levels = unique(results_sepsis$term)),   
    exposures = factor(exposure, levels = exposures))

print(results_sepsis, n = 100)

results_sepsis <- results_sepsis %>% 
  select(term, estimate, conf.low, conf.high)

# Saving data
write_xlsx(results_sepsis, "~/UKL/Malawi ABS Projekt/R Malawi AMS/results_sepsis_corr.xlsx")

# ordering manually in excel ...

# reading df
fp_sepsis_data <- read_excel("~/UKL/Malawi ABS Projekt/R Malawi AMS/Forrest_plot_data_s_corr.xlsx")

#CAVE OR_CI_d (D should be s?)
OR_CI_d <- ifelse(
  fp_sepsis_data$estimate == 1.0,
  "1.0 (Ref)",
  sprintf("%.1f (%.1f–%.1f)", fp_sepsis_data$estimate, fp_sepsis_data$conf.low, fp_sepsis_data$conf.high)
)

FP_sepsis_df <- data.frame(
  Variable = c("Characteristic", fp_sepsis_data$category),
  N = c("N", fp_sepsis_data$n),
  Sepsis = c("Sepsis", fp_sepsis_data$sepsis),
  SepsisPercent = c("%", fp_sepsis_data$sepsis_pct),
  OR_CI = c("OR (95% CI)", OR_CI_d),
  stringsAsFactors = FALSE
)

mean_s <- c(NA, fp_sepsis_data$estimate)
lower_s <- c(NA, fp_sepsis_data$conf.low)
upper_s <- c(NA, fp_sepsis_data$conf.high)

table_text_s <- as.matrix(FP_sepsis_df)
```

**Forrest plot sepsis**

```{r FP_sepsis_plot}

#Plotting
forestplot(
  labeltext = table_text_s,
  mean = mean_s,
  lower = lower_s,
  upper = upper_s,  
  is.summary = c(TRUE, rep(FALSE, nrow(table_text_s) -1)),
  zero = 1,
  xlog = TRUE,
  xlim = c(0.01, 50),
  # xticks = c(0.01, 0.1, 0.5, 1, 2, 5, 10, 20, 50),  
  xticks = c(0.1, 0.5, 1, 2, 5, 10, 20, 50),  
  colgap = unit(3, "mm"),
  col = fpColors(box = "black", lines = "gray", zero = "black"),
  boxsize = 0.2,
  txt_gp = fpTxtGp(
    summary = gpar(cex = 1),
    label = gpar(cex = 0.8),
    ticks = gpar(cex = 0.8)
  ),
  lineheight = unit(0.5, "cm"),
  line.margin = 0.1,
  graph.pos = 5
)
```

**Univariate log regression death**

```{r FP_death_table}
# Creating df for forrest plot - death

# Selecting variables for FP


# Descriptive table
FP_table_1 <- Burns_Sepsis %>%
  select(all_of(exposures), Death_corr) %>%
  pivot_longer(cols = all_of(exposures), names_to = "variable", values_to = "category") %>%
  mutate(variable = factor(variable, levels = exposures)) %>%   #keeping order defined in "exposures"
  group_by(variable, category) %>%
  summarise(
    n = n(),
    deaths = sum(Death_corr == 1, na.rm = TRUE),
    death_pct = round(100 * deaths / n, 1),
    .groups = "drop"
  ) %>% 
  arrange(variable) #keeping order of variabls as given in exposures

print(FP_table_1, n = 100)

#Saving as Excel file
write_xlsx(FP_table_1, "~/UKL/Malawi ABS Projekt/R Malawi AMS/discriptive_death.xlsx")

# OR results table
results <- exposures %>%
  set_names() %>%
  map_df(~ glm(as.formula(paste("Death_corr ~", .x)), data = Burns_Sepsis, 
               family = binomial("logit")) %>% 
           tidy(exponentiate = TRUE, conf.int = TRUE), .id = "exposure") %>%
  filter(term != "(Intercept)")

# keeping order ... 
results <- results %>%
  mutate(
    term = factor(term, levels = unique(results$term)),   
    exposures = factor(exposure, levels = exposures))

print(results, n = 100)

results_death <- results %>% 
  select(term, estimate, conf.low, conf.high)

#Saving as Excel file
write_xlsx(results_death, "~/UKL/Malawi ABS Projekt/R Malawi AMS/results_death.xlsx")


# combined and cleaned tables manually in Excel

# Creating Final DF for Forrest plot
fp_death_data <- read_excel("~/UKL/Malawi ABS Projekt/R Malawi AMS/Forrest_plot_data_d.xlsx")

OR_CI_d <- ifelse(
  fp_death_data$estimate == 1.0,
  "1.0 (Ref)",
  sprintf("%.1f (%.1f–%.1f)", fp_death_data$estimate, fp_death_data$conf.low, fp_death_data$conf.high)
)

FP_death_data <- data.frame(
  Variable = c("Characteristic", fp_death_data$category),
  N = c("N", fp_death_data$n),
  Deaths = c("Deaths", fp_death_data$deaths),
  DeathPercent = c("%", fp_death_data$death_pct),
  OR_CI = c("OR (95% CI)", OR_CI_d),
  stringsAsFactors = FALSE
)

mean <- c(NA, fp_death_data$estimate)
lower <- c(NA, fp_death_data$conf.low)
upper <- c(NA, fp_death_data$conf.high)

table_text_d <- as.matrix(FP_death_data)
```

**Forrest plot death**

```{r FP_death_plot}

# Forrest plot death
forestplot(
  labeltext = table_text_d,
  mean = mean,
  lower = lower,
  upper = upper,
  is.summary = c(TRUE, rep(FALSE, nrow(table_text_d) -1)),
  zero = 1,
  xlog = TRUE,
  xlim = c(0.1, 50),
  xticks = c(0.1, 0.5, 1, 2, 5, 10, 20, 50),  
  colgap = unit(3, "mm"),
  col = fpColors(box = "black", lines = "gray", zero = "black"),
  boxsize = 0.2,
  txt_gp = fpTxtGp(
    summary = gpar(cex = 1),
    label = gpar(cex = 0.8),
    ticks = gpar(cex = 0.8)
    ),
  lineheight = unit(0.5, "cm"),
  line.margin = 0.1,
  graph.pos = 5
)
```

### **Multivariable logistic regression analysis (controlling for TBSA)**

**Death - multivariable logistic regression analysis (controlling for TBSA)**

```{r MV_Regression_Death}
results_death_TBSA <- exposures %>%
  set_names() %>%
  map_df(~ glm(as.formula(paste("Death_corr~TBSA3+", .x)), data = Burns_Sepsis, 
               family = binomial("logit")) %>% 
           tidy(exponentiate = TRUE, conf.int = TRUE), .id = "exposure") %>%
  filter(term != "(Intercept)")

#filtering out TBSA rows
results_death_TBSA <- results_death_TBSA %>%
  filter(!str_detect(term, "TBSA"))

view(results_death_TBSA)
```

**Sepsis - multivariable log regression (controlling for TBSA)**

```{r MV_regression_sepsis}

results_sepsis_TBSA <- exposures %>%
  set_names() %>%
  map_df(~ glm(as.formula(paste("crf5_sepsis_corr~TBSA3+", .x)), 
               data = Burns_Sepsis, 
               family = binomial("logit")) %>% 
           tidy(exponentiate = TRUE, conf.int = TRUE), .id = "exposure") %>%
  filter(term != "(Intercept)")

#filtering out TBSA rows 
results_sepsis_TBSA <- results_sepsis_TBSA %>%
  filter(!str_detect(term, "TBSA"))

view(results_sepsis_TBSA)
```

**Wound infection - multivariable log regression (controlling for TBSA)**

```{r MV_regression_WI}

results_WI_TBSA <- exposures %>%
  set_names() %>%
  map_df(~ glm(as.formula(paste("wound_infection_any~TBSA3+", .x)), 
               data = Burns_Sepsis, 
               family = binomial("logit")) %>% 
           tidy(exponentiate = TRUE, conf.int = TRUE), .id = "exposure") %>%
  filter(term != "(Intercept)")

#filtering out TBSA rows 
results_WI_TBSA <- results_WI_TBSA %>%
  filter(!str_detect(term, "TBSA"))

view(results_WI_TBSA)
```
